---
title: "Biking to Airports is nice "
author:
  - name: Jacob Ford
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
---



# Biking to Airports, it's great! 

For those of you who aren't aware, Toronto has two airports: Pearson, a huge, international hub that is 25 minute express train from downtown; and a smaller, more regionally focused 'City of Toronto' airport called Billy Bishop. Billy-B, as no one I'm aware calls it, is located right off the waterfront in the middle of the city; you could probably hit the CN Tower with a 7 iron. It must be the most urban-located airport in North America, unless I'm missing some evil villain's secret jet strip in Queens. 

A major benefit of this type of locale for an airport is that it allows faithful customers to use multiple means of transit. I used to work in Transportation Planning in the Triangle area in North Carolina. The Raleigh-Durham (RDU) airport sits right in between, well, Raleigh and Durham (Chapel Hill you exist too, I didn't forget). BUT, but, but, there is no easy transit from either city to the airport; even the bus lines are convoluted and push environmentally or budget minded customers to car-share or park-n-ride or whatever the hell a kiss-n-ride is. Not having transit lines extend to airport access for major cities is insane. 

So after we moved to Toronto, I had my eyes on the day I would finally fly out of Silly-Illy-Billy Bishop Airport, as we live close enough that I could hop on a city bike, and cruise over! In short, it was glorious. The views from the Billy Bishop airport are also fantastic, the waiting area, for all three or so gates available, felt more like a nicer university than I went to library. 

I wanted to see if Billy Bishop was the most accessible via bike compared to other major US and Canadian airports. My hypothesis is that it offers access via bike to the greatest number of people, both in absolute and relative terms to the city's population. To answer this, I started pulling isochrone and census data. For this analysis, I'm using mb_isochrone functin from the mapboxapi package, very user friendly and effective. Their 'profile' methods of transportation include driving, walking, cycling, and driving in traffic. A transit map would require different packages, but would be a good follow up to this! 

```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(mapboxapi)
library(mapview)
library(units)  
library(tidycensus)
library(tidyr)

```


## The Competitors

From the American Delegation:
- JFK
- LAX
- O'Hare
- ATL
- SFO
- Dallas Fort Worth
- PHL
- Seattle
- Denver

From the Canadian Delegation:
- Toronto Pearson
- Silly Billy
- Montreal
- Vancouver




```{r message=FALSE, warning=FALSE, include=FALSE}
# Define airports with their coordinates (US and Canada), updating Billy Bishop's coordinates
airports <- data.frame(
  name = c("JFK", 
           "LAX", 
           "Chicago O'Hare",
           "Toronto Pearson", 
           "Vancouver", 
           "ATL",
           "SFO",
           "Montreal",
           "Calgary",
           "Dallas/Fort Worth",
           "Toronto Billy Bishop",
           "Philadelphia",
           "Seattle",
           "Denver"),
  lon = c(-73.7781, -118.4085, -87.9048, 
          -79.6306, -123.184, -84.4277, 
          -122.379, -73.7415, -114.0134,
          -97.0403, -79.39769267068745, -75.24114,
          -122.3088, -104.6737),
  lat = c(40.6413, 33.9416, 41.9742, 
          43.6777, 49.1947, 33.6407, 
          37.6213, 45.4577, 51.1315,
          32.8968, 43.63409886780112, 39.8744,
          47.4502, 39.8561)
)

# Load the sf library if not already loaded
library(sf)

# Convert to sf object
airports <- st_as_sf(airports, coords = c("lon", "lat"), crs = 4326)


```


```{r message=FALSE, warning=FALSE, include=FALSE}

access_token = "pk.eyJ1IjoiamFrZWZvcmQ5MjciLCJhIjoiY2xrd3A2aHhzMDFzODNlbHNmdW02cTZlaCJ9.2-bZutoqPYKahlOtsTJszg"

isos_60 <- airports %>%
  mb_isochrone(time = 60, profile = "cycling", id_column = "name",
               access_token = access_token) %>%
  mutate(isos_60 = geometry)

isos_45 <- airports %>%
  mb_isochrone(time = 45, profile = "cycling", id_column = "name",
               access_token = access_token)%>%
  mutate(isos_45 = geometry)

isos_30 <- airports %>%
  mb_isochrone(time = 30, profile = "cycling", id_column = "name",
               access_token = access_token)%>%
  mutate(isos_30 = geometry)


```

```{r message=FALSE, warning=FALSE, include=FALSE}
# combine all the isos dfs
isos <- cbind(isos_60, isos_45, isos_30)

canadian_airs <- c("Vancouver", "Toronto", "Calgary", "Montreal")

canadian_isos <- isos %>%
  filter(grepl(paste(canadian_airs, collapse = "|"), id)) 


canadian_isos_30 <- isos_30 %>%
  filter(grepl(paste(canadian_airs, collapse = "|"), id)) 

canadian_isos_45 <- isos_45 %>%
  filter(grepl(paste(canadian_airs, collapse = "|"), id)) 

canadian_isos_60 <- isos_60 %>%
  filter(grepl(paste(canadian_airs, collapse = "|"), id)) 

us_isos <- isos %>%
  filter(!grepl(paste(canadian_airs, collapse = "|"), id)) 

us_isos_30 <- isos_30 %>%
  filter(!grepl(paste(canadian_airs, collapse = "|"), id)) 

us_isos_45 <- isos_45 %>%
  filter(!grepl(paste(canadian_airs, collapse = "|"), id)) 

us_isos_60 <- isos_60 %>%
  filter(!grepl(paste(canadian_airs, collapse = "|"), id)) 

```

# Isochrone Madness

Transportation Modeling 101: **Isochrones**: a travel distance map, taking into account existing networks. Usually measured in time, it uses aggregated and average travel distance (sometimes by time of day, AM vs PM) to plot out how far one could travel from a given point. These ink-blot looking items on the below map represent a 60 minute isochrone from each of the major airports we consider. You may have to zoom in to see these. 

For the analysis below, we'll consider isochrones at 30, 45, and 60 minute intervals. The smaller the isochrone, the less distance one would be able to travel from the given airport. 



```{r echo=FALSE, message=FALSE, warning=FALSE}
mapview(isos)
```


What you'll notice is that most airports are placed relatively far from the city they serve. This is sensible given the space these facilities require for safe operations and storage. If you're familiar with Toronto, you'll see the Billy Bishop, JFK and Seattle isochrones are huge; whereas poor Denver is lil. The below graph shows the total area covered by each of the three isochrone sets. At first, it seems my hypothesis is ruined; not only is Silly Billy among the lowest in total area covered, but it's not even better than Pearson (which doesn't exactly make sense if you live here). However, it does track that Denver's airport is the lowest, given it is located approximately 15 light years from downtown Denver and likely does not have bike lanes to/fro that many people would be able to take. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Assuming 'isos' is your data frame of isochrones with geometry
area_compares <- isos %>%
  mutate(
      area_30 = st_area(isos$isos_30) * 3.86102e-7,
      area_45 = st_area(isos$isos_45) * 3.86102e-7,
      area_60 = st_area(isos$isos_60) * 3.86102e-7
     
      ) %>%  # Convert square meters to square miles
  st_drop_geometry() %>%
  group_by(id) %>%
  arrange(desc(area_30))  

# sort by largest area
# Load necessary library


# Strip units if necessary (you can skip this step if you want to keep units in plots)
area_compares$area_30 <- as.numeric(units::drop_units(area_compares$area_30))
area_compares$area_45 <- as.numeric(units::drop_units(area_compares$area_45))
area_compares$area_60 <- as.numeric(units::drop_units(area_compares$area_60))


# Assuming your dataframe is named df
temp_long <- area_compares %>%
  mutate(Airport = id) %>%
  pivot_longer(
    cols = starts_with("area"), 
    names_to = "Duration", 
    values_to = "Area"
  )

# Create the bar chart
ggplot(temp_long, aes(x = Airport, y = Area, fill = Duration)) +
  geom_bar(stat = "identity", position = position_dodge()) +  # position_dodge to separate the bars
  labs(title = "Area at Different Isochrones",
       x = "Airport",
       y = "Area",
       fill = "Area (sq meters)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels for better readability





# # Assuming 'area_compares' already has the areas calculated and stripped of units
# 
# # Create a bar chart for the 30-minute isochrone area
# ggplot(area_compares, aes(x = reorder(id, -area_30), y = area_30)) +
#   geom_bar(stat = "identity", color = "black") +  # Use identity to use the actual area values
#   theme_minimal() +  # Optional: cleaner theme
#   labs(title = "30 Minute Biking Isochrone Area",
#        x = "",
#        y = "Area (square miles)",
#        fill = "Isochrone ID") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels for better readability

# # Create a bar chart for the 45-minute isochrone area
# ggplot(area_compares, aes(x = reorder(id, -area_45), y = area_45)) +
#   geom_bar(stat = "identity", color = "black") +
#   theme_minimal() +
#   labs(title = "45 Minute Biking Isochrone Area",
#        x = "",
#        y = "Area (square miles)",
#        fill = "Isochrone ID") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# # Create a bar chart for the 60-minute isochrone area
# ggplot(area_compares, aes(x = reorder(id, -area_60), y = area_60)) +
#   geom_bar(stat = "identity", color = "black") +
#   theme_minimal() +
#   labs(title = "60 Minute Biking Isochrone Area",
#        x = "",
#        y = "Area (square miles)",
#        fill = "Isochrone ID") +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


# Show me the People

Who cares about area coverage? Well you should have, but not anymore! Let's translate those isochrones into how many people are covered within those areas. We'll use public US and Canadian census data to overlay loaded tracts with populations, then do a simple geospatial overlay to translate how many people are within the isochrone. This will show: 

1. How many people are able to bike to airports in a given X minute interval 
2. Give a better comparison metric between airports that take into account the concentration and distribution of actual humans, instead of area. 



```{r message=FALSE, warning=FALSE, include=FALSE}
state_list <- c("WA", "NY", "GA", "CA", "TX",
                "PA", "IL", "CO")

# use get_acs to get population data 
# Get population data for all US counties
us_data <- get_acs(
  geography = "county",
  variables = "B01003_001",
  year = 2022,
  state = state_list,
  geometry = TRUE
)

## 30 minute 

us_isos_30 <- st_transform(us_isos_30, st_crs(us_data))


# Perform spatial intersection
intersection <- st_intersection(us_isos_30, us_data)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

us_data$tract_area <- st_area(us_data)
temp_us_data <- us_data %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_us_data)

final_temp_30 <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * estimate) %>%
  group_by(Airport = id) %>%
  summarize(Population_30 = sum(temp_pop))


## 45 minute

us_isos_45 <- st_transform(us_isos_45, st_crs(us_data))


# Perform spatial intersection
intersection <- st_intersection(us_isos_45, us_data)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

us_data$tract_area <- st_area(us_data)
temp_us_data <- us_data %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_us_data)

final_temp_45 <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * estimate) %>%
  group_by(Airport = id) %>%
  summarize(Population_45 = sum(temp_pop))



## 60 minute

  
us_isos_60 <- st_transform(us_isos_60, st_crs(us_data))


# Perform spatial intersection
intersection <- st_intersection(us_isos_60, us_data)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

us_data$tract_area <- st_area(us_data)
temp_us_data <- us_data %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_us_data)

final_temp_60 <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * estimate) %>%
  group_by(Airport = id) %>%
  summarize(Population_60 = sum(temp_pop))

final_us <- cbind(final_temp_30, final_temp_45, final_temp_60)

```



```{r message=FALSE, warning=FALSE, include=FALSE}

# Canada

library(cancensus)

options(cancensus.api_key = "CensusMapper_a46eeccdea7abd430a378f4f43ac2c21")
toronto_tracts <- get_census(dataset = 'CA21', 
                          regions = list(CMA = c("35535", "537")),  # Codes for Toronto CMA and Hamilton CMA
                          vectors = c("v_CA21_1"), 
                          level = 'CT', 
                          use_cache = FALSE, 
                          geo_format = 'sf',
                          quiet = TRUE)

van_tracts <- get_census(dataset = 'CA21', 
                          regions = list(CMA = c("59933")),  # Codes for Toronto CMA and Hamilton CMA
                          vectors = c("v_CA21_1"), 
                          level = 'CT', 
                          use_cache = FALSE, 
                          geo_format = 'sf',
                          quiet = TRUE)



montreal_tracts <- get_census(dataset = 'CA21', 
                          regions = list(CMA = "24462"),  # Code for Montréal CMA
                          vectors = c("v_CA21_1"),  # Ensure this vector code corresponds to the data you want
                          level = 'CT', 
                          use_cache = FALSE, 
                          geo_format = 'sf',  # Returns spatial data frame
                          quiet = TRUE)
```






```{r message=FALSE, warning=FALSE, include=FALSE}

## 30 minute 

canadian_isos_30 <- st_transform(canadian_isos_30, st_crs(toronto_tracts))

#####--30 minutes--#####

# Toronto

intersection <- st_intersection(canadian_isos_30, toronto_tracts)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

toronto_tracts$tract_area <- st_area(toronto_tracts)
temp_toronto_data <- toronto_tracts %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_toronto_data)

final_temp_30_toronto <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * Population) %>%
  group_by(Airport = id) %>%
  summarize(Population_30 = sum(temp_pop))

# Montreal

intersection <- st_intersection(canadian_isos_30, montreal_tracts)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

montreal_tracts$tract_area <- st_area(montreal_tracts)
temp_montreal_data <- montreal_tracts %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_montreal_data)

final_temp_30_montreal <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * Population) %>%
  group_by(Airport = id) %>%
  summarize(Population_30 = sum(temp_pop))

# Vancouver

intersection <- st_intersection(canadian_isos_30, van_tracts)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

van_tracts$tract_area <- st_area(van_tracts)
temp_van_data <- van_tracts %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_van_data)

final_temp_30_van <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * Population) %>%
  group_by(Airport = id) %>%
  summarize(Population_30 = sum(temp_pop))

final_temp_30_candian <- rbind(final_temp_30_toronto, final_temp_30_montreal, final_temp_30_van)


#####--45 minutes--#####

# Toronto

intersection <- st_intersection(canadian_isos_45, toronto_tracts)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

toronto_tracts$tract_area <- st_area(toronto_tracts)
temp_toronto_data <- toronto_tracts %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_toronto_data)

final_temp_45_toronto <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * Population) %>%
  group_by(Airport = id) %>%
  summarize(Population_45 = sum(temp_pop))

# Montreal

intersection <- st_intersection(canadian_isos_45, montreal_tracts)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

montreal_tracts$tract_area <- st_area(montreal_tracts)
temp_montreal_data <- montreal_tracts %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_montreal_data)

final_temp_45_montreal <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * Population) %>%
  group_by(Airport = id) %>%
  summarize(Population_45 = sum(temp_pop))

# Vancouver

intersection <- st_intersection(canadian_isos_45, van_tracts)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

van_tracts$tract_area <- st_area(van_tracts)
temp_van_data <- van_tracts %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_van_data)

final_temp_45_van <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * Population) %>%
  group_by(Airport = id) %>%
  summarize(Population_45 = sum(temp_pop))

final_temp_45_candian <- rbind(final_temp_45_toronto, final_temp_45_montreal, final_temp_45_van)



#####--60 minutes--#####

  
# Toronto

intersection <- st_intersection(canadian_isos_60, toronto_tracts)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

toronto_tracts$tract_area <- st_area(toronto_tracts)
temp_toronto_data <- toronto_tracts %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_toronto_data)

final_temp_60_toronto <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * Population) %>%
  group_by(Airport = id) %>%
  summarize(Population_60 = sum(temp_pop))

# Montreal

intersection <- st_intersection(canadian_isos_60, montreal_tracts)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

montreal_tracts$tract_area <- st_area(montreal_tracts)
temp_montreal_data <- montreal_tracts %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_montreal_data)

final_temp_60_montreal <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * Population) %>%
  group_by(Airport = id) %>%
  summarize(Population_60 = sum(temp_pop))

# Vancouver

intersection <- st_intersection(canadian_isos_60, van_tracts)

# Calculate the area of each intersection
intersection$intersect_area <- st_area(intersection)

van_tracts$tract_area <- st_area(van_tracts)
temp_van_data <- van_tracts %>% st_drop_geometry()

temp_intersection <- intersection %>%
  left_join(temp_van_data)

final_temp_60_van <- temp_intersection %>%
  st_drop_geometry() %>%
  mutate(temp_pop = intersect_area/tract_area * Population) %>%
  group_by(Airport = id) %>%
  summarize(Population_60 = sum(temp_pop))

final_temp_60_candian <- rbind(final_temp_60_toronto, final_temp_60_montreal, final_temp_60_van)


final_canadian <- cbind(final_temp_30_candian, final_temp_45_candian, final_temp_60_candian)


```

The results are below: The bad news, Silly Billy is not number 1... but it's number 2! JFK leads the way, which fine, whatever, likely makes sense there are 4x as many people in NYC than Toronto. BUT! I will claim a modicum of vindication given over 500k people live within a 30 minute bike ride to Silly Billy, and over 1.5 million live within an hour bike ride. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

final <- rbind(final_us, final_canadian)

# Load necessary libraries
library(ggplot2)
library(tidyr)
library(dplyr)

# Assuming your dataframe is named df
df_long <- final %>%
  pivot_longer(
    cols = starts_with("Population"), 
    names_to = "Duration", 
    values_to = "Population"
  )

# Create the bar chart
ggplot(df_long, aes(x = Airport, y = Population, fill = Duration)) +
  geom_bar(stat = "identity", position = position_dodge()) +  # position_dodge to separate the bars
  labs(title = "Airport Populations at Different Durations",
       x = "Airport",
       y = "Population",
       fill = "Pop within X minute Bike") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x labels for better readability

```
# Conclusion

Biking to the airport is freeing, gets your body loose before you're crammed into an aisle seat, is a carbon-free means of transit, is significantly cheaper than car-sharing or parking fees, and is overall awesome! If you ever get the chance to do it, do it! 



```{r message=FALSE, warning=FALSE, include=FALSE}
# Percent of MSA Population

metro_areas <- data.frame(
  MSA = c("ATL", 
          "Chicago", 
          "Dallas",
          "Denver", 
          "JFK", 
          "LAX", 
          "Philadelphia", 
          "Seattle", 
          "SFO", 
          "Montreal", 
          "Toronto Billy Bishop",
          "Toronto Pearson", 
          "Vancouver"), 
  Population = c(6307261, 9459000, 6488000, 2897000, 
                 19200000, 13000000, 6100000, 4000000, 4700000,
                 
                 # to do: update canadian MSA pop estimates
                 15000000, 3400000, 3400000, 1500000
                 )
)


```


