---
title: "Good Eatin's"
description: |
  The hidden gems I didn't want to share
author:
  - name: Jacob Ford
    url: https://jford-dchcmpo.github.io/Blog/posts/2021-12-30-Redistricting-and-Racial-Targetting/
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
---

Click on a location, get my 'don't miss' item(s) from each. 

August 2022 update: Avoiding the doldrums of local reporting by inventive new gimics, the News & Observer's March Madness 2021 [burger bracket](https://www.newsobserver.com/living/food-drink/article255290876.html), that actually took place in the fall of 2021 - bestowed upon the brassiere [Neuse River Brewing](https://www.neuseriverbrewing.com/) the grand title, the tsar of taste, the sultan of sizzle, the best of the burgers. We had high expectations, they were somehow exceeded. 10/10 must try. 

December update: Krill, you must go to Krill if you're in Durham. 



```{r setup, include=FALSE}
library(dplyr)
library(leaflet)
library(sf) 

library(readxl)
eats_data <- read_excel("Eat_Data.xlsx")


```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(RColorBrewer)
pal <- colorFactor(
  palette = 'Dark2',
  domain = eats_data$Genre
)


eats_map <- leaflet(data=eats_data) %>%
 # addTiles() %>%
 addProviderTiles("CartoDB.Positron") %>%
  addCircles(lat = eats_data$Latitutde, lng=eats_data$Longitude,
             popup=paste("Name: ", eats_data$Name, "<br>",
                          "Genre: ", eats_data$Genre, "<br>",
                          "Don't you dare miss: ", eats_data$`Don't Miss:`),
             color = ~pal(Genre),
             weight=10,
             fillOpacity = 2) %>%
  addLegend("bottomleft",
            pal=pal,
            values=~Genre) 
  


eats_map
```




```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

## Searchable Map
bbq <- eats_data %>%
  filter(Genre=="BBQ")

eats_map <- leaflet() %>%
 # addTiles() %>%
 addProviderTiles("CartoDB.Positron") %>%
  addCircles(lat = eats_data$Latitutde, lng=eats_data$Longitude,
             popup=paste("Name: ", eats_data$Name, "<br>",
                          "Genre: ", eats_data$Genre, "<br>",
                          "Don't you dare miss: ", eats_data$`Don't Miss:`),
           #  color = ~pal(Genre),
             group="All",
             data=eats_data) %>%
  addCircles(lat = eats_data$Latitutde, lng=eats_data$Longitude,
             popup=paste("Name: ", eats_data$Name, "<br>",
                          "Genre: ", eats_data$Genre, "<br>",
                          "Don't you dare miss: ", eats_data$`Don't Miss:`),
            #color = ~pal(Genre),
             group="BBQ",
             data=bbq) %>%
  
  # addLegend("bottomleft",
  #           pal=pal,
  #           values=~Genre)  %>%
    addLayersControl(overlayGroups = c("BBQ", "All"))

  


eats_map
```


